---
title: 'Bayesian power analysis: Part III'
subtitle: "All the world isn't a Gauss."
output:
  github_document
---

## tl;dr

So far we've covered Bayesian power simulations from both a null hypothesis orientation and a parameter width perspective. In both instances, we kept things simple and stayed with Gaussian (i.e., normally distributed) data. But not all data follow that form, so it might do us well to expand our skill set a bit. In this post, we'll cover how we might perform power simulations with count and binary data. For the count data, we'll use the Poisson likelihood. For the binary, we'll use the binomial.

## The Poisson distribution is handy for counts.

In the social sciences, count data arise when we ask questions like:

* How many sexual partners have you had?
* How many pets do you have at home?
* How many cigarettes did you smoke, yesterday?

The values these data will take are discrete [^1] in that you've either slept with 9 or 10 people, but definitely not 9.5. The values cannot go below zero in that even if you quit smoking cold turkey 15 years ago and have been a health nut since, you still could not have smoked -3 cigarettes, yesterday. Zero is as low as it goes. 

The canonical distribution for data of this type--non-zero integers--is the Poisson. It's named after the French mathematician Siméon Denis Poisson, [who had quite the confident stare in his youth](https://upload.wikimedia.org/wikipedia/commons/e/e8/E._Marcellot_Siméon-Denis_Poisson_1804.jpg). The Poisson distribution has one parameter, $\lambda$, which controls both its mean and variance. Although the numbers the Poisson describes are counts, the $\lambda$ parameter does not need to be an integer. For example, here's the plot of 1000 draws from a Poisson for which $\lambda = 3.2$.

```{r, fig.width = 4, fig.height = 2.5, warning = F, message = F}
library(tidyverse)

theme_set(theme_gray() + theme(panel.grid = element_blank()))

tibble(x = rpois(n = 1e3, lambda = 3.2)) %>% 
  mutate(x = factor(x)) %>% 
  
  ggplot(aes(x = x)) +
  geom_bar()
```

In case you missed it, the key function for generating those data was `rpois()`. I'm not going to go into a full-blown tutorial on the Poisson distribution or on count regression. For more thorough introductions, check out Atkins et al's [*A tutorial on count regression and zero-altered count models for longitudinal substance use data*](https://www.ncbi.nlm.nih.gov/pmc/articles/PMC3513584/pdf/nihms396181.pdf), chapters 9 through 11 in McElreath's [*Statistical Rethinking*](https://xcelab.net/rm/statistical-rethinking/), or, if you really want to dive in, Agresti's [*Foundations of Linear and Generalized Linear Models*](https://www.wiley.com/en-us/Foundations+of+Linear+and+Generalized+Linear+Models-p-9781118730034).

For our power example, let's say you were interested in drinking. Using data from [the National Epidemiologic Survey on Alcohol and Related Conditions](https://pubs.niaaa.nih.gov/publications/AA70/AA70.htm), Christopher Ingraham presented [a data visualization](https://www.washingtonpost.com/news/wonk/wp/2014/09/25/think-you-drink-a-lot-this-chart-will-tell-you/?utm_term=.b81599bbbe25) of the average number of alcoholic drinks American adults consume, per week. By decile, the numbers were: 

1. 0.00
2. 0.00 
3. 0.00
4. 0.02
5. 0.14 
6. 0.63
7. 2.17
8. 6.25
9. 15.28
10. 73.85

Let's say you wanted to run a study where you planned on comparing two demographic groups by their weekly drinking levels. Let's further say you suspected one of those groups drank like the folks American adults in the 7^th^ decile and the other drank like American adults in the 8^th^. We'll call them low and high drinkers, respectively. For convenience, let's further presume you'll be able to recruit equal numbers of participants from both groups. The question for our power analysis is to determine how many you'd need per group to detect reliable differences. Using $n = 50$ as a starting point, here’s what the data for our hypothetical groups might look like.

```{r, fig.width = 4, fig.height = 4}
mu_7 <- 2.17
mu_8 <- 6.25

n <- 50

set.seed(3)

d <-
  tibble(low  = rpois(n = n, lambda = mu_7),
         high = rpois(n = n, lambda = mu_8)) %>% 
  gather(group, count) 

d %>%
  mutate(count = factor(count)) %>% 
  
  ggplot(aes(x = count)) +
  geom_bar() +
  facet_wrap(~group, ncol = 1)
```

This will be our primary data type. Our next step is to determine how to express our research question as a regression model. Like with our two-group Gaussian models, we can predict counts in terms of an intercept (i.e., standing for the expected value on the reference group) and slope (i.e., standing for the expected difference between the reference group and the comparison group). If we coded our two groups by a `high` variable for which 0 stood for low drinkers and 1 stood for high drinkers, the basic model would follow the form

$$
\begin{align*}
\text{drinks_per_week}_i & \sim \text{Poisson}(\lambda_i) \\
\text{log} (\lambda_i)   & = \beta_0 + \beta_1 \text{high}_i.
\end{align*}
$$

Here's how to set the data up for that model.

```{r}
d <-
  d %>% 
  mutate(high = ifelse(group == "low", 0, 1))
```

  
If you were attending closely to our model formula, you noticed we ran into a detail. Count regression, such as with the Poisson likelihood, tends to use the log link. *Why?* you ask. Recall that counts need to be 0 and above. Same deal for our $\lambda$ parameter. In order to make sure our models don't yield silly estimates for $\lambda$, like -2 or something, we typically use the log link. You don't have to, of course. [The world is your playground](https://i.kym-cdn.com/entries/icons/original/000/008/342/ihave.jpg). But this is the method most of your colleagues are likely to use and it's the one I suggest you use until you have compelling reasons to do otherwise.

So then since we're now fitting a model with a log link, it might seem challenging to pick good priors. As a place to start, we can use the `brms::get_prior()` function to see the **brms** defaults.

```{r, warning = F, message = F}
library(brms)

get_prior(data = d,
          family = poisson,
          count ~ 0 + intercept + high)
```

Hopefully two things popped out. First, there's no prior of `class = sigma`. Since the Poisson distribution only has one parameter $\lambda$, we don't need to set a prior for $\sigma$. Our model won't have one. Second, because we're continuing to use the `0 + intercept` syntax for our model intercept, both our intercept and slope are of prior `class = b` and those currently have default flat priors with **brms**. To be sure, flat priors aren't the best. But maybe if this was your first time playing around with a Poisson model, default flat priors might seem like a safe place to start. [Feel free to disagree](https://xkcd.com/386/). In the meantime, here's how to fit that default Poisson model with `brms::brm()`.

```{r fit1, cache = T, message = F, warning = F, results = "hide"}
fit1 <-
  brm(data = d,
      family = poisson,
      count ~ 0 + intercept + high,
      seed = 3)
```

```{r}
print(fit1)
```

Since we used the log link, our model results are in the log metric, too. If you'd like them in the metric of the data, you'd work directly with the poster samples and exponentiate. 

```{r}
post <- 
  posterior_samples(fit1) %>% 
  mutate(`beta_0 (i.e., low)`                       = exp(b_intercept),
         `beta_1 (i.e., difference score for high)` = exp(b_high))
```

We can then just summarize our parameters of interest.

```{r}
post %>% 
  select(starts_with("beta_")) %>% 
  gather() %>% 
  group_by(key) %>% 
  summarise(mean  = mean(value),
            lower = quantile(value, prob = .025),
            upper = quantile(value, prob = .975))
```

For the sake of simulation, it'll be easier if we press on with evaluating the parameters on the log metric, though. If you're working within a null-hypothesis oriented power paradigm, you'll be happy to know zero is still the number to beat for evaluating our 95% intervals for $\beta_1$, even when that parameter is in the log metric. Here it is, again.

```{r, warning = F, message = F}
library(broom)

tidy(fit1, prob = .95) %>% 
  filter(term == "b_high")
```

So our first fit suggests we're on good footing to run a quick power simulation holding $n = 50$. As in the prior blog posts, our lives will be simpler if we set up a custom simulation function. Since we'll be using it to simulate the data and fit the model in one step, let's call it `sim_data_fit()`.

```{r}
sim_data_fit <- function(seed, n) {
  
  n <- n
  
  set.seed(seed)
  
  d <-
    tibble(high  = rep(0:1, each = n),
           count = c(rpois(n = n, lambda = mu_7),
                     rpois(n = n, lambda = mu_8)))
  
  update(fit1,
         newdata = d,
         seed = seed) %>% 
  tidy(prob = .95) %>% 
  filter(term == "b_high") %>% 
  select(lower:upper)
  
}
```

Here's the simulation for a simple 100 iterations.

```{r sim1, cache = T, warning = F, message = F, results = "hide"}
sim1 <-
  tibble(seed = 1:100) %>% 
  mutate(ci = map(seed, sim_data_fit, n = 50)) %>% 
  unnest()
```

That went quick--less than 3 minutes on my old laptop. Here's what those 100 $\beta_1$ intervals look like in bulk.

```{r, fig.width = 8, fig.height = 3}
sim1 %>% 
  
  ggplot(aes(x = seed, ymin = lower, ymax = upper)) +
  geom_hline(yintercept = 0, color = "white") +
  geom_linerange() +
  labs(x = "seed (i.e., simulation index)",
       y = expression(beta[1]))
```

None of them are anywhere near the null value 0. So it appears we're well above .8 power to reject the typical $H_0$ with $n = 50$. Here's the distribution of their widths.

```{r, fig.width = 4, fig.height = 2.5}
sim1 %>% 
  mutate(width = upper - lower) %>% 
  
  ggplot(aes(x = width)) +
  geom_histogram(binwidth = .01)
```

What if we wanted a mean width of 0.25 on the log scale? We might try the simulation with $n = 150$.

```{r sim2, cache = T, warning = F, message = F, results = "hide"}
sim2 <-
  tibble(seed = 1:100) %>% 
  mutate(ci = map(seed, sim_data_fit, n = 150)) %>% 
  unnest()
```

Here we'll summarize the widths both in terms of their mean and what proportion were smaller than 0.25.

```{r}
sim2 %>% 
  mutate(width = upper - lower) %>% 
  summarise(`mean width` = mean(width),
            `below 0.25` = mean(width < 0.25))
```

If we wanted to focus on the mean, we did pretty good. Perhaps set the $n = 155$ and simulate a full 1000+ iterations for a serious power analysis. But if we wanted to make the stricter criteria of all below 0.25, we'd need to up the $n$ quite a bit more. And of course, once you have a little experience working with Poisson models, you might do the power simulations with more ambitious priors. For example, if your count values are lower than like 1000, there's a good chance a `normal(0, 6)` prior on your $\beta$ parameters will be nearly flat within the reasonable neighborhoods of the parameter space.

```{r, fig.width = 5, fig.height = 2, eval = F, echo = F}
# here's a quick visualization of why that's so
tibble(sd = 0:6,
       `sd * 2` = sd * 2) %>% 
  gather() %>% 
  
  ggplot(aes(x = value, y = exp(value))) +
  geom_point() +
  facet_wrap(~key, scales = "free")
```

Now you've got a sense of how to work with the Poisson likelihood, it's time to move on to the binomial.

## The binomial is handy for binary data.

Binary data are even weirder than counts. They typically only take on two values: 0 and 1. Sometimes 0 is a stand-in for "no" and 1 for "yes" (e.g., *Are you an expert in Bayesian power analysis?* For me that would be `0`). You can also have data of this kind if you asked people whether they'd like to choose option A or B. With those kinds of data, you might code A as 0 and B as 1. Binomial data also often stand in for trials where 0 = "fail" and 1 = "success." For example, if you answered "Yes" to the question *Are all data normally distributed?* we'd mark your answer down as a `0`.

Though 0s and 1s are popular, sometimes binomial data appear in their aggregated form. Let's say I gave you 10 algebra questions and you got 7 of them right. Here's one way to encode those data.

```{r}
n <- 10
z <- 7

rep(0:1, times = c(n - z, z))
```

In that example, `n` stood for the total number of trials and `z` was the number you got correct (i.e., the number of times we encoded your response as a 1). A more compact way to encode that data is with two columns, one for `z` and the other for `n`.

```{r}
tibble(z = z,
       n = n)
```

So then if you gave those same 10 questions to four of your friends, we could encode the results like this.

```{r}
set.seed(3)

tibble(id = letters[1:5],
       z  = rpois(n = 5, lambda = 5),
       n  = n)
```

If you're `b`, it appears you're the smart one in the group.

Anyway, whether working with binomial or aggregated binomial data, we're interested in the probability a given trial will be 1. 

### Logistic regression with binary data.

Taking binary data as a starting point, given $d$ data that includes a variable $y$ where the value in the $i^\text{th}$ row is a 0 or a 1, we'd like to know $p(y_i = 1 | d)$. The binomial distribution will help us get that estimate for $p$. We'll do so within the context of a logistic regression model following the form

$$
\begin{align*}
y_i                & \sim \text{Binomial} (n = 1, p_i) \\
\text{logit} (p_i) & = \beta_0,
\end{align*}
$$

were the logit function id defined as the log odds

$$
\text{logit} (p_i) = \text{log} \bigg (\frac{p_i}{1 - p_i} \bigg),
$$

and thus

$$
\text{log} \bigg (\frac{p_i}{1 - p_i} \bigg) = \beta_0.
$$

In those formulas, $\beta_0$ is the intercept. In a binomial model with no predictors [^2], the intercept $\beta_0$ is just the estimate for $p$, but in the log-odds metric. So yes, similar to the Poisson models, we typically use a link function with our binomial models. Instead of the log link, we use the logit because constrains the posterior for $p$ to values between 0 and 1. Just as the null value for a probability is .5, the null value for the parameters within a logistic regression model is typically 0. 

As above with the Poisson, I'm not going to go into a full-blown tutorial on the binomial distribution or on logistic regression. For more thorough introductions, check out chapters 9 through 10 in McElreath's [*Statistical Rethinking*](https://xcelab.net/rm/statistical-rethinking/) or Agresti's [*Foundations of Linear and Generalized Linear Models*](https://www.wiley.com/en-us/Foundations+of+Linear+and+Generalized+Linear+Models-p-9781118730034).

Time to simulate some data. Let's say we'd like to estimate the probability someone will hit a ball in a baseball game. Nowadays, batting averages for professional baseball players tend around .25 (see [here](http://www.baseball-almanac.com/hitting/hibavg4.shtml)). So if we wanted to simulate 50 at-bats, we might do so like this.

```{r}
set.seed(3)

d <- tibble(y = rbinom(n = 50, size = 1, prob = .25))

str(d)
```

Here are what those data look like in a bar plot.

```{r, fig.width = 3, fig.height = 2}
d %>% 
  mutate(y = factor(y)) %>% 
  
  ggplot(aes(x = y)) +
  geom_bar()
```

Here's the **brms** default for our intercept-only logistic regression model.

```{r}
get_prior(data = d, 
          family = binomial,
          y | trials(1) ~ 1)
```

That's a really liberal prior. We might be a little more gutsy and put a more skeptical `normal(0, 2)` prior on that intercept. With the context of our logit link, that still puts a 95% probability that the $p$ is between .02 and .98, which is almost the entire parameter space. Here's how to fit the model with the `brm()` function.

```{r, eval = F, echo = F}
# proof of how permissive that `normal(0, 2)` prior is
inv_logit_scaled(c(-4, 4))
```


```{r fit2, cache = T, message = F, warning = F, results = "hide"}
fit2 <-
  brm(data = d, 
      family = binomial,
      y | trials(1) ~ 1,
      prior(normal(0, 2), class = Intercept),
      seed = 3)
```

In the `brm()` formula syntax, including a `|` bar on the left side of a formula indicates we have extra supplementary information about our criterion variable. In this case, that information is that each `y` value corresponds to a single trial (i.e., `trials(1)`), which itself corresponds to the $n = 1$ portion of the statistical formula, above. Here are the results.

```{r}
print(fit2)
```

Remember that that intercept is on the scale of the logit link, the log odds. We can transform it with the `brms::inv_logit_scaled()` function.

```{r}
fixef(fit2)["Intercept", 1] %>% 
  inv_logit_scaled()
```

If we want the full posterior distribution, we'll need to work with the posterior draws themselves.

```{r, fig.width = 4, fig.height = 2.25}
posterior_samples(fit2) %>% 
  transmute(p = inv_logit_scaled(b_Intercept)) %>% 
  
  ggplot(aes(x = p)) +
  geom_density(fill = "grey25", size = 0) +
  scale_x_continuous("probability of a hit", limits = c(0, 1)) +
  scale_y_continuous(NULL, breaks = NULL)
```

Looks like the null hypothesis of $p = .5$ is not credible for this simulation. If we'd like the posterior median and percentile-based 95% intervals, we might use the `median_qi()` function from the handy [**tidybayes** package](https://mjskay.github.io/tidybayes/index.html).

```{r, warning = F, message = F}
library(tidybayes)

posterior_samples(fit2) %>% 
  transmute(p = inv_logit_scaled(b_Intercept)) %>% 
  median_qi()
```

Yep, .5 was not within those intervals. Let's see what happens when we do a mini power analysis with 100 iterations. First we set up our simulation function.

```{r}
sim_data_fit <- function(seed, n_player) {
  
  n_trials <- 1
  prob_hit <- .25
  
  set.seed(seed)
  
  d <- tibble(y = rbinom(n    = n_player, 
                         size = n_trials, 
                         prob = prob_hit))
  
  update(fit2,
         newdata = d,
         seed = seed) %>% 
  posterior_samples() %>% 
  transmute(p = inv_logit_scaled(b_Intercept)) %>% 
  median_qi() %>% 
    select(.lower:.upper)
  
}
```

Simulate.

```{r sim3, cache = T, warning = F, message = F, results = "hide"}
sim3 <-
  tibble(seed = 1:100) %>% 
  mutate(ci = map(seed, sim_data_fit, n_player = 50)) %>% 
  unnest()
```

You might plot the intervals.

```{r, fig.width = 8, fig.height = 3}
sim3 %>% 
  ggplot(aes(x = seed, ymin = .lower, ymax = .upper)) +
  geom_hline(yintercept = c(.25, .5), color = "white") +
  geom_linerange() +
  xlab("seed (i.e., simulation index)") +
  scale_y_continuous("probability of hitting the ball", limits = c(0, 1))
```

Like one of my old coworkers used to say: *Purtier 'n a hog!* Here we'll summarize the results both in terms of their conventional power, their mean width, and the proportion of widths more narrow than .25. *Why .25?* I don't know. Without a substantively-informed alternative, it's as good a criterion as any.

```{r}
sim3 %>% 
  mutate(width = .upper - .lower) %>% 
  summarise(`conventional power` = mean(.upper < .5),
            `mean width`         = mean(width),
            `width below .25`    = mean(width < .25))
```

Depending on your study needs, you'd adjust your sample size accordingly, do a mini simulation or two first, and then follow up with a proper 1000+ power simulation.

I should point out that whereas we evaluated the power of the Poisson model with the parameters on the scale of the link function, we evaluated the power for our logistic regression model after transforming the intercept back into the probability metric. Both methods are fine. The way you run your power simulation should be based on how you want to interpret and report your results.

We should also acknowledge that this was our first example of a power simulation that wasn't based on some group comparison. Comparing groups is fine and normal and important. And it's also the case that power matters for more than group-based analyses. Our simulation-based approach to Bayesian power analyses is fine for both.

### Aggregated binomial regression.

It's no more difficult to simulate and work with aggregated binomial data. But since the mechanics for `brms::brm()` and thus the down-the-road simulation setup are a little different, we should practice. With our new setup, we'll consider a new example. Since .25 is the typical batting average, it might better sense to define the null hypothesis like this:

$$H_0 \text{: } p = .25$$

Consider a case where we had some intervention where we expected a new batting average of .35. How many trials would we need, then, to either reject $H_0$ or at least estimate $p$ with a satisfactory degree of precision? Here's what the statistical formula for the implied aggregated binomial model might look like:

$$
\begin{align*}
y_i                & \sim \text{Binomial} (n, p_i) \\
\text{logit} (p_i) & = \beta_0.
\end{align*}
$$

The big change is we no longer defined $n$ as 1. Let’s say we wanted our aggregated binomial data set to contain the summary statistics for $n = 100$ trials. Here’s what that might look like.

```{r}
n_trials <- 100
prob_hit <- .35

set.seed(3)

d <- tibble(n_trials = n_trials,
            y = rbinom(n    = 1, 
                       size = n_trials, 
                       prob = prob_hit))

d
```

Now we have two columns. The first, `n_trials`, indicates how many cases or trials we're summarizing. The second, `y`, indicates how many successes/1s/hits we might expect given $p = .35$. This is the aggregated binomial equivalent of if we had a 100 row vector composed of 32 1s and 68 0s.

Now, before we discuss fitting the model with **brms**, let's talk priors. Since we've updated our definition of $H_0$, it might make sense to update the prior for $\beta_0$. As it turns out, setting that prior to `normal(-1, 0.5)` puts the posterior mode at about .25 on the probability space, but with fairly wide 95% intervals ranging from about .12 to .5. Though centered on our updated null value, this prior is still quite permissive given our hypothesized $p = .35$.

```{r, fig.width = 3, fig.height = 2.25, echo = F, eval = F}
# here's the evidence for that claim
set.seed(1)
proof <-
  tibble(y = rnorm(1e6, -1, .5)) %>% 
  mutate(y = inv_logit_scaled(y))

proof %>% 
  ggplot(aes(x = y, y = 0)) +
  geom_halfeyeh(point_interval = mode_qi, .width = c(.5, .95)) +
  scale_x_continuous("hit probability", limits = c(0, 1)) +
  scale_y_continuous(NULL, breaks = NULL) +
  labs(subtitle = "This is the consequence of the\nnormal(-1, 0.5) prior after converting it\nback to the probability metric.")

proof %>% 
  mode_qi()

rm(proof)
```

To fit an aggregated binomial model with the `brm()` function, we augment the `<criterion> | trials()` syntax where the value that goes in `trials()` is either a fixed number or variable in the data indexing $n$. Our approach will be the latter.

```{r fit3, cache = T, message = F, warning = F, results = "hide"}
fit3 <-
  brm(data = d, 
      family = binomial,
      y | trials(n_trials) ~ 1,
      prior(normal(-1, 0.5), class = Intercept),
      seed = 3)
```

Inspect the summary.

```{r}
print(fit3)
```

After a transformation, here’s what that looks like in a plot.

```{r, fig.width = 4, fig.height = 2}
posterior_samples(fit3) %>% 
  transmute(p = inv_logit_scaled(b_Intercept)) %>% 
  
  ggplot(aes(x = p, y = 0)) +
  geom_halfeyeh(.width = c(.5, .95)) +
  scale_x_continuous("probability of a hit", limits = c(0, 1)) +
  scale_y_continuous(NULL, breaks = NULL)
```

Based on a single simulation, it looks like $n = 100$ won’t quite be enough to reject $H_0 \text{: } p = .25$ with a conventional 2-sided 95% interval. But it does look like we're in the ballpark and that our basic data + model setup will work for a larger-scale simulation. Here’s an example of how you might update our custom simulation function.

```{r}
sim_data_fit <- function(seed, n_trials) {
  
  prob_hit <- .35
  
  set.seed(seed)
  
  d <- tibble(y = rbinom(n    = 1, 
                         size = n_trials, 
                         prob = prob_hit),
              n_trials = n_trials)
  
  update(fit3,
         newdata = d,
         seed = seed) %>% 
  posterior_samples() %>% 
  transmute(p = inv_logit_scaled(b_Intercept)) %>% 
  median_qi() %>% 
    select(.lower:.upper)
  
}
```

Simulate, this time trying out $n = 120$.

```{r sim4, cache = T, warning = F, message = F, results = "hide"}
sim4 <-
  tibble(seed = 1:100) %>% 
  mutate(ci = map(seed, sim_data_fit, n_trials = 120)) %>% 
  unnest()
```

Plot the intervals.

```{r, fig.width = 8, fig.height = 3}
sim4 %>% 
  ggplot(aes(x = seed, ymin = .lower, ymax = .upper)) +
  geom_hline(yintercept = c(.25, .35), color = "white") +
  geom_linerange() +
  xlab("seed (i.e., simulation index)") +
  scale_y_continuous("probability of hitting the ball",
                     limits = c(0, 1), breaks = c(0, .25, .35, 1))
```

Overall, those intervals look pretty good. They're fairly narrow and are hovering around the data generating $p = .35$. But it seems many are still crossing the .25 threshold. Let's see the results of a formal summary.

```{r}
sim4 %>% 
  mutate(width = .upper - .lower) %>% 
  summarise(`conventional power` = mean(.lower > .25),
            `mean width`         = mean(width),
            `width below .2`     = mean(width < .2))
```

All widths were more narrow than .2 and the mean width was about .16. In the abstract that might seem reasonably precise. But we're still not precise enough to reject $H_0$ with a conventional power level. Depending on your needs, adjust the $n$ accordingly and simulate again.

## Session info

```{r}
sessionInfo()
```

## Footnotes

[^1]: Yes, one can smoke half a cigarette or drink 1/3 of a drink. Ideally, we'd have the exact amount of nicotine in your blood at a given moment and over time and the same for the amount of alcohol in your system relative to your blood volume and such. But in practice, substance use researchers just don't tend to have access to data of that quality. Instead, we're typically stuck with simple counts. And I look forward to the day the right team of engineers, computer scientists, and substance use researchers (and whoever else I forgot to mention) release the cheap, non-invasive technology we need to passively measure these things. Until then: *How many standard servings of alcohol did you drink, last night?*

[^2]: In case this is all new to you and you and you had the question in your mind: Yes, you can add predictors to the logistic regression model. Say we had a model with two predictors, $x_1$ and $x_2$. Our statistical model would then follow the form $\text{logit} (p_i) = \beta_0 + \beta_1 x_{1i} + \beta_2 x_{2i}$.

